---
title: "What Comes Next Monte Carlo Simulation on Estimates"
author: "Michael Gordon"
date: "29/09/2020"
output: html_document
---

Using the three point estimates provided to us by our team of experts we have used the beta-PERT distribution to generate simulations of the project tasks, we are assuming that the estimates provided are fairly accurate which is why we have chosen the beta-PERT distribution as it more closely resembles a normal distribution and places more emphasis around the “most likely” value, we have configured the lambda parameter with the default value of 4 to maintain a close resemblance to a normal distribution (RiskAMP 2020).

The beta-PERT Distribution | RiskAMP (2020). Available at: https://www.riskamp.com/beta-pert (Accessed: 30 September 2020).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(gridExtra)

theme_set(theme_bw())
theme_update( plot.title = element_text(hjust=0.5)
            , plot.subtitle = element_text(hjust=0.5)
            )

estimates <- read_csv('data/estimates_task_estimates.csv')
salaries <- read_csv('data/estimates_salaries_2.csv')

# PERT estimator courtesy of https://www.riskamp.com/beta-pert
pert_dist_estimator <- function( n, x.min, x.max, x.mode, lambda = 4 ){

    if( x.min > x.max || x.mode > x.max || x.mode < x.min ) stop( "invalid parameters" );

    x.range <- x.max - x.min;
    if( x.range == 0 ) return( rep( x.min, n ));

    mu <- ( x.min + x.max + lambda * x.mode ) / ( lambda + 2 );

    # special case if mu == mode
    if( mu == x.mode ){
        v <- ( lambda / 2 ) + 1
    }
    else {
        v <- (( mu - x.min ) * ( 2 * x.mode - x.min - x.max )) /
            (( x.mode - mu ) * ( x.max - x.min ));
    }

    w <- ( v * ( x.max - mu )) / ( mu - x.min );
    return ( rbeta( n, v, w ) * x.range + x.min );
}
```


```{r}
n <- 10000
set.seed(42)
data_engineer_estimates <- estimates %>% filter(Implementor == 'Data Engineer/Data Scientist' & Type != 'Documentation')
full_stack_estimates <- estimates %>% filter(Implementor == 'Full Stack developer' & Type != 'Documentation')

data_engineer_documentation_estimates <- estimates %>% filter(Implementor == 'Data Engineer/Data Scientist' & Type == 'Documentation')
full_stack_documentation_estimates <- estimates %>% filter(Implementor == 'Full Stack developer' & Type == 'Documentation')

monte_carlo_model <- function(data_df){
  simulation_results_df <- as.data.frame(matrix(nrow=n,ncol=nrow(data_df)))
  
  for(i in 1:nrow(data_df)){
    results = pert_dist_estimator(n, data_df$Minimum[i], data_df$Maximum[i], data_df$MostLikely[i])
    simulation_results_df[,i] <- results
  }
  
  return(simulation_results_df)
}

data_engineer_simulations <- monte_carlo_model(data_engineer_estimates)
full_stack_simulations <- monte_carlo_model(full_stack_estimates)
data_engineer_documentation_simulations <- monte_carlo_model(data_engineer_documentation_estimates)
full_stack_documentation_simulations <- monte_carlo_model(full_stack_documentation_estimates)

sim_sum <- function(data_df){
  result <- rep(0, nrow(data_df))
  for(i in 1:ncol(data_df)){
    result = result + data_df[,i]
  }
  return(result)
}

data_engineer_total_dev_time <- data.frame(time = sim_sum(data_engineer_simulations))
data_engineer_total_doc_time <- data.frame(time = sim_sum(data_engineer_documentation_simulations))
full_stack_total_dev_time <- data.frame(time = sim_sum(full_stack_simulations))
full_stack_total_doc_time <- data.frame(time = sim_sum(full_stack_documentation_simulations))
```

```{r, fig.width=10}
plot_simulation <- function(simulation_times, bin_size, x_labels = c("Completed on day", "Completed by day"), lower_bound_point = .05, upper_bound_point = .95){
  
  e = ecdf(simulation_times$time)
  upper_x_point <- quantile(e,c(upper_bound_point),type=7)
  lower_x_point <- quantile(e,c(lower_bound_point),type=7)
  
  grid.arrange(
    ggplot(simulation_times, aes(x = time)) + 
      geom_histogram(aes(y=..density..),      
                     binwidth=bin_size,
                     colour="black", fill="white") +
      geom_vline(aes(xintercept = lower_x_point), linetype="dotted", color = "red") +
      geom_vline(aes(xintercept = upper_x_point), linetype="dotted", color = "red") +
      scale_y_continuous(labels = scales::comma) +
      scale_x_continuous(labels = scales::comma) +
      xlab(x_labels[1]) +
      geom_density(alpha=.2, fill="#FF6666") +
      ggtitle("Probability Distribution Curve")
    
    ,ggplot(simulation_times, aes(time)) +
      stat_ecdf(geom = 'line') +

      geom_point(aes(y = upper_bound_point, x = upper_x_point), colour = 'red') +
      geom_point(aes(y = lower_bound_point, x = lower_x_point), colour = 'red') +
      
      scale_y_continuous(labels = scales::percent, name = "Probability") +
      scale_x_continuous(labels = scales::comma) +
      xlab(x_labels[2]) +
      ggtitle("Cumulative Probability Distribution Curve")
    , ncol = 2
  )
}
```

### Data Modelling/Engineering development time in days

```{r, fig.width=10}
plot_simulation(data_engineer_total_dev_time, 1)
```

```{r}
quantile(data_engineer_total_dev_time$time,c(0.05,0.95),type=7)
```

### Data Medelling/Engineering documentation time in days

```{r, fig.width=10}
plot_simulation(data_engineer_total_doc_time, 1)
```

```{r}
quantile(data_engineer_total_doc_time$time,c(0.05,0.95),type=7)
```

### Application Development time in days

```{r, fig.width=10}
plot_simulation(full_stack_total_dev_time, 1)
```

```{r}
quantile(full_stack_total_dev_time$time,c(0.05,0.95),type=7)
```

### Application Documentation time in days

```{r, fig.width=10}
plot_simulation(full_stack_total_doc_time, 1)
```

```{r}
quantile(full_stack_total_doc_time$time,c(0.05,0.95),type=7)
```

### Total project time in days

We are planning for the application development and data engineering/machine learning work to be done in parallel, for elements of the application development which are dependent on the NLP model a stub web service will be developed which will return dummy results.

```{r, fig.width=10}
full_stack_total_time <- data.frame(time = full_stack_total_doc_time$time + full_stack_total_dev_time$time)
data_engineer_total_time <- data.frame(time = data_engineer_total_doc_time$time + data_engineer_total_dev_time$time)

total_duration <- data.frame(time = pmax(full_stack_total_time$time, data_engineer_total_time$time))
plot_simulation(total_duration, 1)
```

```{r}
quantile(total_duration$time,c(0.05,0.95),type=7)
```

### Salaries 

We have salary estimates for junior, senior and lead data scientist/data engineer for the ETL, ML and NLP work, and full stack developer for the the application development work, we will use these estimates to perform a monte carlo simulation, and adjust for anticipated efficiency of each skill level. Based on this simulation we are 90% confident that the development effort will cost between 45,000 & 73,000.

```{r}

efficiency <- c(
  c(0.45,0.95),
  c(0.15,0.65),
  c(0.05,0.55)
)

# our salary estimates are annual, but our time estimates are in days, so we need to convert
# the salaries to daily rates, salary conversion formula: https://staff.uow.edu.au/personnel/salary/UOW020681.html
days = 260.893
salary_full_stack <- c(
  (salaries %>% filter(Type == 'Full Stack developer') %>% select(Junior))[[1]] / days,
  (salaries %>% filter(Type == 'Full Stack developer') %>% select(Senior))[[1]] / days,
  (salaries %>% filter(Type == 'Full Stack developer') %>% select(Lead))[[1]] / days
)

salary_data_engineer <- c(
  (salaries %>% filter(Type == 'Data Engineer/Data Scientist') %>% select(Junior))[[1]] / days,
  (salaries %>% filter(Type == 'Data Engineer/Data Scientist') %>% select(Senior))[[1]] / days,
  (salaries %>% filter(Type == 'Data Engineer/Data Scientist') %>% select(Lead))[[1]] / days
)

simulate_cost <- function(salary, simulations){
  cost_estimates <- data.frame(
    sim_id = numeric(),
    cost = numeric()
  )
  
  set.seed(42)
  for(si in 1:n){
    costs = numeric(ncol(simulations))
    for(i in 1:ncol(simulations)){
      
      level <- sample.int(3, 1, replace = T)
      adjusted_sim <- quantile(simulations[,i],efficiency[level],type=7)
      
      costs[i] <- sample(adjusted_sim, 1) * salary[level]
    }
    
    cost_estimates <- cost_estimates %>% union(
      data.frame(sim_id = si, cost = sum(costs))
    )
  }
  return(cost_estimates)
}

cost_estimates_full_stack <- simulate_cost(salary_full_stack, full_stack_simulations)
cost_estimates_data_engineer <- simulate_cost(salary_data_engineer, data_engineer_simulations)
```

#### Total Cost of development based on time and salary estimates

```{r, fig.width=10}

total_costs <- data.frame(time = cost_estimates_data_engineer$cost + cost_estimates_full_stack$cost)

plot_simulation(total_costs, 500, c("Total Cost", "Total Cost"))
```

```{r}
quantile(total_costs$time,c(0.05,0.95),type=7)
```
